Equalizer 2.0 --\\Convergence of a Parallel Rendering Framework

### 2018-02-15 paperIEEE.tex
Address FIXMEs
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@132 db7c848b-7755-4df2-b763-8796797afa52

\textsf{Equalizer} itself has received significant attention within the research
community. Various algorithms to improve the parallel rendering performance have
been proposed: compression and region of interest during
compositing~\cite{MEP:10}; load-balancing resources for multi-display
installations~\cite{EEP:11}; asynchronous compositing and NUMA
optimizations~\cite{EBAHMP:12}; as well as work queueing~\cite{SPEP:16}. \FIXME{Why ; instead of , ?}
Additionally, complex large scale and out-of-core multiresolution
rendering approaches have been parallelized and implemented with
Equalizer~\cite{GMBP:10, GEMPG:13}, demonstrating the feasibility of
the framework to be used with complex rendering algorithms and 3D model
representations.

@@

Sort-first (\fig{fig:loadeq}) and sort-last load balancing are the most obvious
optimizations for these parallel rendering modes. Our load equalizers are fully
transparent for application developers; that is, they use a reactive approach
based on past rendering times. This assumes a reasonable frame-to-frame
coherency. \FIXME{Shouldn't that be frame-to-frame {\em coherence}?} Equalizer implements two different algorithms, a
\textsf{load\_equalizer} and a \textsf{tree\_equalizer}. The result section
provides some evidence on the strengths and weaknesses of both algorithms.

@@

\FIXME{Sentence incomplete.}
Asynchronous compositing pipelines rendering with compositing operations, by
executing the image readback, network transfer and image assembly from threads
running in parallel to the rendering threads. In \cite{EBAHMP:12}, we provide
the details of the implementation and experimental data showing an improvement
of the rendering performance of over 25\% for large node counts.

@@

\begin{compactdesc}
\item[Static] The object is not versioned nor buffered. The instance data is
  serialized whenever a new slave instance is mapped. No additional data is
  stored.
\item[Instance] The object is versioned and buffered. The instance and delta
  data are identical; that is, only instance data is serialized. Previous
  instance data is saved to be able to map old versions.
\item[Delta] The object is versioned and buffered. The delta data is typically
  smaller than the instance data. The delta data is transmitted to slave
  instances for synchronization. Previous instance and delta data is saved to be
  able to map and sync old versions.
\item[Unbuffered] The object is versioned and unbuffered. No data is stored, and
  no previous versions can be mapped. \FIXME{Redundant.} No old versions can be mapped and no
  additional data is stored.
\end{compactdesc}

Serialization is facilitated using output or input streams, which abstract the
data transmission and are used like an \textsf{std::stream}. The data streams
implement efficient buffering and compression, and automatically select the best
connection for data transport. Custom data type serializers can be implemented
by providing the appropriate serialization functions. No pointers should be
directly transmitted through the data streams. For pointers, the corresponding
object is typically a distributed object as well, and its UUID and version are
transmitted in place of a pointer.

@@

\begin{compactitem}
\item The master instance of the object generates new versions for all
  slaves. These versions are continuous. It is possible to commit on slave
  instances, but special care has to be taken to handle possible
  conflicts.
\item Slave instance versions can only be advanced; that is, \textsf{sync(
  version)} with a version smaller than the current version will fail.
\item Newly mapped slave instances are mapped to the oldest available
  version by default \FIXME{Is this correct?}, or to the version specified when calling
  \textsf{mapObject}.
\end{compactitem}

@@

\textsf{Collage} implements a few generic distributed objects which are used by
\textsf{Equalizer} and other applicatons. A barrier is a distributed barrier
primitive used for software swap barriers in \textsf{Equalizer}
(\sref{sec:swap}) \FIXME{Using the word barrier three times in one short sentence}. Its implementation follows a simple master-slave approach,
which has shown to by sufficient for this use case. The used distributed data structures are single producer,
multiple consumer FIFO queues. To hide network latencies, consumers prefetch stored
items into a local queue. Queues are used for tile and chunk compounds
(\sref{sec:tile}).

@@

Omegalib \cite{Omegalib} is a software framework built on top of Equalizer that
facilitates application development for hybrid reality environments such as the
Cave 2. Hybrid reality enviroments aim to create a seamless 2D/3D environment
that supports both information-rich analysis (traditionally done on tiled
display wall) as well as virtual reality simulation exploration (traditionally
done in VR systems) at a resolution matching human visual acuity. Omegalib
supports dynamic reconfigurability of the display environment, so that areas of
the display can be interactively allocated to 2D or 3D workspaces as needed. It
makes it possible to have multiple immersive applications running on a
cluster-controlled display system, have different input sources dynamically
routed to applications, and have rendering results optionally redirected to a
distributed compositing manager. Omegalib supports pluggable front-ends, to
simplify the integration of third-party libraries like OpenGL \FIXME{But if it's based on Equalizer then OpenGL should already be part of the backend?}, OpenSceneGraph,
and the Visualization Toolkit (VTK).

@@

A critical piece for data distribution performance are the characteristics of
the data compression algorithm. Our microbenchmark compresses a set of binary
files \FIXME{What are these files? Fig. 7 only speaks of "generic" binary data...}, precalculating the speed and compression ratio of the various engines.
\fig{fCompressor} shows the compression and decompression speed in gigabyte per
second as well as the size of the compressed data relative to the uncompressed
data. The ZSTD$x$ engines use the ZStandard compression library at compression
level $x$. The measurements were performed on a single, isolated node.

@@

\FIXME{Sentence incomplete.}
used the same data sets as in the previous section. We ran the benchmark on up
to eight physical nodes, that is, after eight processes nodes start to run two
processes per node, which share CPU, memory and network interface bandwidth.

>>>>

\textsf{Equalizer} itself has received significant attention within the research
community. Various algorithms to improve the parallel rendering performance have
been proposed: compression and region of interest during
compositing~\cite{MEP:10}, load-balancing resources for multi-display
installations~\cite{EEP:11}, asynchronous compositing and NUMA
optimizations~\cite{EBAHMP:12}, as well as work queueing~\cite{SPEP:16}.
Additionally, complex large scale and out-of-core multiresolution
rendering approaches have been parallelized and implemented with
Equalizer~\cite{GMBP:10, GEMPG:13}, demonstrating the feasibility of
the framework to be used with complex rendering algorithms and 3D model
representations.

@@

Sort-first (\fig{fig:loadeq}) and sort-last load balancing are the most obvious
optimizations for these parallel rendering modes. Our load equalizers are fully
transparent for application developers; that is, they use a reactive approach
based on past rendering times. This assumes a reasonable frame-to-frame
coherence. Equalizer implements two different algorithms, a
\textsf{load\_equalizer} and a \textsf{tree\_equalizer}. The result section
provides some evidence on the strengths and weaknesses of both algorithms.

@@

Asynchronous compositing is pipelining the rendering with compositing operations, by
executing the image readback, network transfer and image assembly from threads
running in parallel to the rendering threads. In \cite{EBAHMP:12}, we provide
the details of the implementation and experimental data showing an improvement
of the rendering performance of over 25\% for large node counts.

@@

\begin{compactdesc}
\item[Static] The object is not versioned nor buffered. The instance data is
  serialized whenever a new slave instance is mapped. No additional data is
  stored.
\item[Instance] The object is versioned and buffered. The instance and delta
  data are identical; that is, only instance data is serialized. Previous
  instance data is saved to be able to map old versions.
\item[Delta] The object is versioned and buffered. The delta data is typically
  smaller than the instance data. The delta data is transmitted to slave
  instances for synchronization. Previous instance and delta data is saved to be
  able to map and sync old versions.
\item[Unbuffered] The object is versioned and unbuffered. No data is stored, and
  no previous versions can be mapped.
\end{compactdesc}

Serialization is facilitated using output or input streams, which abstract the
data transmission and are used like a \textsf{std::stream}. The data streams
implement efficient buffering and compression, and automatically select the best
connection for data transport. Custom data type serializers can be implemented
by providing the appropriate serialization functions. No pointers should be
directly transmitted through the data streams. For pointers, the corresponding
object is typically a distributed object as well, and its UUID and version are
transmitted in place of a pointer.

@@

\begin{compactitem}
\item The master instance of the object generates new versions for all
  slaves. These versions are continuous. It is possible to commit on slave
  instances, but special care has to be taken to handle possible
  conflicts.
\item Slave instance versions can only be advanced; that is, \textsf{sync(
  version)} with a version smaller than the current version will fail.
\item Newly mapped slave instances are mapped to the oldest available
  version by default, or to the version specified when calling
  \textsf{mapObject}.
\end{compactitem}

@@

\textsf{Collage} implements a few generic distributed objects which are used by
\textsf{Equalizer} and other applications. A barrier is a distributed barrier
primitive used for software swap synchronization in \textsf{Equalizer}
(\sref{sec:swap}). Its implementation follows a simple master-slave approach,
which has shown to by sufficient for this use case. Queues are distrbuted, single producer,
multiple consumer FIFO queues. To hide network latencies, consumers prefetch
items into a local queue. Queues are used for tile and chunk compounds
(\sref{sec:tile}).

@@

Omegalib \cite{Omegalib} is a software framework built on top of Equalizer that
facilitates application development for hybrid reality environments such as the
Cave 2. Hybrid reality enviroments aim to create a seamless 2D/3D environment
that supports both information-rich analysis (traditionally done on tiled
display wall) as well as virtual reality simulation exploration (traditionally
done in VR systems) at a resolution matching human visual acuity. Omegalib
supports dynamic reconfigurability of the display environment, so that areas of
the display can be interactively allocated to 2D or 3D workspaces as needed. It
makes it possible to have multiple immersive applications running on a
cluster-controlled display system, have different input sources dynamically
routed to applications, and have rendering results optionally redirected to a
distributed compositing manager. Omegalib supports pluggable front-ends, to
simplify the integration of third-party libraries like OpenGL, OpenSceneGraph,
and the Visualization Toolkit (VTK).

@@

A critical piece for data distribution performance are the characteristics of
the data compression algorithm. Our microbenchmark compresses a set of binary
files, precalculating the speed and compression ratio of the various engines.
\fig{fCompressor} shows the compression and decompression speed in gigabyte per
second as well as the size of the compressed data relative to the uncompressed
data. The ZSTD$x$ engines use the ZStandard compression library at compression
level $x$. The measurements were performed on a single, isolated node.

@@

We used the same data sets as in the previous section, and ran the benchmark on up
to eight physical nodes, that is, after eight processes nodes start to run two
processes per node, which share CPU, memory and network interface bandwidth.
